\chapter{Selected References}

The following references are the official or primary documentation anchors used for this manual. They are not a substitute for reading the full API specifications when relying on edge-case behavior.

\begin{thebibliography}{99}
\bibitem{OpenJDK25} OpenJDK. \emph{JDK 25 Project}. \url{https://openjdk.org/projects/jdk/25/}
\bibitem{JDK25Docs} Oracle. \emph{JDK 25 Documentation Home}. \url{https://docs.oracle.com/en/java/javase/25/}
\bibitem{JavaDesktopModule} Oracle. \emph{Module java.desktop, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/module-summary.html}
\bibitem{SwingPackage} Oracle. \emph{Package javax.swing, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/package-summary.html}
\bibitem{JComponent} Oracle. \emph{JComponent, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/JComponent.html}
\bibitem{RepaintManager} Oracle. \emph{RepaintManager, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/RepaintManager.html}
\bibitem{DebugGraphics} Oracle. \emph{DebugGraphics, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/DebugGraphics.html}
\bibitem{JViewport} Oracle. \emph{JViewport, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/JViewport.html}
\bibitem{EventQueue} Oracle. \emph{EventQueue, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/EventQueue.html}
\bibitem{SwingUtilities} Oracle. \emph{SwingUtilities, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/SwingUtilities.html}
\bibitem{SwingWorker} Oracle. \emph{SwingWorker, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/SwingWorker.html}
\bibitem{Thread} Oracle. \emph{Thread, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.base/java/lang/Thread.html}
\bibitem{Graphics2D} Oracle. \emph{Graphics2D, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/Graphics2D.html}
\bibitem{MultiResolutionImage} Oracle. \emph{MultiResolutionImage, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/image/MultiResolutionImage.html}
\bibitem{JTable} Oracle. \emph{JTable, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/JTable.html}
\bibitem{RowSorter} Oracle. \emph{RowSorter, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/RowSorter.html}
\bibitem{TableRowSorter} Oracle. \emph{TableRowSorter, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/table/TableRowSorter.html}
\bibitem{JTree} Oracle. \emph{JTree, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/JTree.html}
\bibitem{Action} Oracle. \emph{Action, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/Action.html}
\bibitem{JLayer} Oracle. \emph{JLayer, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/JLayer.html}
\bibitem{LayerUI} Oracle. \emph{LayerUI, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/plaf/LayerUI.html}
\bibitem{UIManager} Oracle. \emph{UIManager, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/UIManager.html}
\bibitem{ComponentUI} Oracle. \emph{ComponentUI, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/plaf/ComponentUI.html}
\bibitem{SwingTextPackage} Oracle. \emph{Package javax.swing.text, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/text/package-summary.html}
\bibitem{TransferHandler} Oracle. \emph{TransferHandler, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/swing/TransferHandler.html}
\bibitem{DataFlavor} Oracle. \emph{DataFlavor, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.datatransfer/java/awt/datatransfer/DataFlavor.html}
\bibitem{AccessibleContext} Oracle. \emph{AccessibleContext, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/javax/accessibility/AccessibleContext.html}
\bibitem{InputMethodRequests} Oracle. \emph{InputMethodRequests, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/im/InputMethodRequests.html}
\bibitem{JLinkModule} Oracle. \emph{Module jdk.jlink, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/jdk.jlink/module-summary.html}
\bibitem{JPackage} Oracle. \emph{The jpackage Command, JDK 25 Tool Specifications}. \url{https://docs.oracle.com/en/java/javase/25/docs/specs/man/jpackage.html}
\bibitem{Desktop} Oracle. \emph{Desktop, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/Desktop.html}
\bibitem{Taskbar} Oracle. \emph{Taskbar, Java SE 25 API}. \url{https://docs.oracle.com/en/java/javase/25/docs/api/java.desktop/java/awt/Taskbar.html}
\end{thebibliography}
